Accessing the COVID-19 Mortality Rate in the US: While It Varies Across Different Income Levels, the Impact of Political Preference is Unignorable

### 2023-11-17 outputs/paper/paper.qmd
Revise the paper

library(tidyverse)
library(here)
library(modelsummary)
library(gridExtra)
library(usmap)
library(knitr)
library(kableExtra)
library(purrr)
library(ggplot2)

@@

As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS) [@acs], The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu], and Fox News [@fox]. Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county.

@@

After cleaning each data set, I merge the five data sets by county and state. Based on the merged data set, I calculate the infection and death number per 1000 residents in each county. Therefore, each row in the merged data represents a US county and provides demographic information such as the total population. In addition, the data also contains the voting pattern in each county. The @tbl-summary_variables summarizes all the important variables in this paper.

@@

In each OLS regression model, the dependent variable will consistently be the COVID-19 death rate for each county. Given that death rates can be influenced by many factors, including demographics and economics, and given our interest in the potential impact of political party affiliation from the last Federal Election, I will establish two sets of models. The first set will not incorporate political preferences, while the second set will. Within each of these sets, there will be three models, each focusing exclusively on one category of predictors. Additionally, there will be a comprehensive model encompassing all predictors and a best-fit model determined by $R^2_{adj}$ and $RMSE$. In total, this approach will yield ten models, including the two best-fit models. To decide which model is the best, I will evaluate them using testing data to see which one has a lower testing $RMSE$.

@@

## Expected MSE and Bias-Variance Trade-off

The rationale for partitioning the data into training and testing sets is to validate our model. That said, by training a model using training data and then evaluating it using test data, we can determine which model produces the lowest expected MSE and better performance.

$$
\mathbb{E}\left[(Y-\hat{f}(x))^2 \mid X=x\right]=\operatorname{Var}(\hat{f}(x))+(\mathbb{E}[\hat{f}(x)]-f(x))^2+\sigma^2
$$
The above equation shows the expected MSE, which underscores the bias-variance trade-off. It tells us the trade-off between the model's complexity and the interpretability. An intricate model may increase the bias, but its variance will be higher if we vary the data, and vice-versa.

Hence, by evaluating our models using the testing data, we can know whether we should fit a relatively simple or complicated model.

@@

## Comparing the testing error

The above two models are fitted using the training data. Comparing the training RMSE from @tbl-all_models1 and @tbl-all_models2, we can find that the model containing the political preference has a lower training RMSE. As introduced in the Methods part, to test the prediction performance of the models, I will calculate the test error for the two models to see which is better.

```{r}
#| echo: false
#| message: false
#| label: tbl-test_error
#| tbl-cap: The testing error of the above two models

calculate_testing_error <- function(model, testing_data, outcome) {
  # Predict the outcomes on the testing data
  predictions <- predict(model, newdata = testing_data)

  # Compute the residuals
  residuals <- predictions - testing_data[[outcome]]

  # Compute the MSE
  mse <- mean(residuals^2)

  return(mse)
}

data.frame(model = c("Model 1", "Model 2"),
           error = c(calculate_testing_error(best1, testing_data, "death_rate"),
                     calculate_testing_error(best2, testing_data, "death_rate"))) %>% 
  rename(`Model` = model,
         `Testinf Error` = error) %>% 
  kable(booktabs = TRUE)
```

@tbl-test_error shows the testing error of the above two models. We can observe that the errors for the two models are 1.769733 and 1.675969. Therefore, the model containing the political preferences has a lower testing MSE; it has better predictions than the another one.

>>>>

library(tidyverse)
library(here)
library(modelsummary)
library(gridExtra)
library(usmap)
library(knitr)
library(kableExtra)
library(purrr)
library(ggplot2)
library(boot)

@@

As introduced in the introduction, there are three primary sources of data in the data: the American Community Survey (ACS) [@acs], The Center for Systems Science and Engineering (CSSE) at John Hopkins University [@jhu], and Fox News [@fox]. Each source provides distinct data dimensions that contribute to predicting the COVID-19 death rate in each US county. The @tbl-summary_variables introduces all the important variables in this paper.

@@

After cleaning each data set, I merge the five data sets by county and state. Based on the merged data set, I calculate the infection and death number per 1000 residents in each county. Therefore, each row in the merged data represents a US county and provides demographic information such as the total population. In addition, the data also contains the voting pattern in each county.

@@

In each OLS regression model, the dependent variable will consistently be the COVID-19 death rate for each county. Given that death rates can be influenced by many factors, including demographics and economics, and given our interest in the potential impact of political party affiliation from the last Federal Election, I will establish two sets of models. The first set will not incorporate political preferences, while the second set will. Within each of these sets, there will be three models, each focusing exclusively on one category of predictors. Additionally, there will be a comprehensive model encompassing all predictors and a best-fit model determined by $R^2_{adj}$ and $RMSE$. In total, this approach will yield ten models, including the two best-fit models. To decide which model is the best, I will perform the 10-fold Cross-Validation to

@@

## 10-fold Cross-Validation

The way to compare the two models is to calculate the error by performing the 10-fold Cross-Validation. 10-fold Cross-Validation means to to randomly split the data to 10 roughly equal-sized folds and the first one is treated as the validation set. Then fit the model on the remaining folds and compute the MSE in the first fold. Repeat the process for ten times and the CV error is computed as the averaging of these values. That is:

$$
CV_{(10)}=\frac{1}{10}\sum_{i=1}^{10}MSE_i
$$
Hence, we can compare the CV error for the two models and the model will a lower CV error will perform better.

@@

## 10 Fold Cross-Validation Error

@tbl-cv_error shows the 10-fold Cross-Validation error of the above two models. We can see that the model containing the political preference has a smaller error. Therefore, it has a better performance than another one and we should consider the county's political preference when we predict the COVID mortality rate of each county.

```{r}
#| echo: false
#| message: false
#| label: tbl-cv_error
#| tbl-cap: The testing error of the above two models

set.seed(0)
glm.fit <- glm(best1, data =training_data)
cv_error_1 <- cv.glm(training_data, glm.fit, K=10)$delta[1]
glm.fit <- glm(best2, data = training_data)
cv_error_2 <- cv.glm(training_data, glm.fit, K=10)$delta[1]

data.frame(model = c("Model 1", "Model 2"),
           error = c(cv_error_1,
                     cv_error_2)) %>% 
  rename(`Model` = model,
         `CV Error` = error) %>% 
  kable(booktabs = TRUE)
```
